\textbf{International Transport costs:\\New Findings from modeling additive costs}\\Online Appendix (Not for Publication)

### 2018-12-17 redaction/TC_new.tex
changements typos graphiques

\title{\textsc{International Transport costs:\\New Findings from modeling additive costs}\thanks{We are grateful to Olger G\"{o}rg, S\'{e}bastien Jean, Vincent Vicard for their insightful comments. The paper has greatly benefited from remarks from the participants at ETSG 17\textsuperscript{th} (Helsinki), DEGIT XXII (Paris), seminars at the Universites of California-Irvine, Paris-Dauphine, Paris-I, Nancy and the the Kiel Institute for the World Economy.
Finally, any remaining errors are ours.
This paper was partly funded by an University grant from Lille-I University.
It features an online appendix containing additional results and available on the authors' websites.}}

\author{Guillaume \textsc{Daudin}\thanks{%
Universit\'{e} Paris-Dauphine \& PSL University, IRD, LEDa, UMR 225, DIAL, 75016 PARIS, FRANCE \newline
Sciences Po, Observatoire Fran\c{c}ais des Conjonctures \'{E}conomiques (OFCE), 75014 PARIS, FRANCE \newline
; email: \url{guillaume.daudin@dauphine.fr}}  \qquad J\'{e}r\^{o}me \textsc{H\'{e}ricourt} \thanks{Universit\'{e} de Lille - LEM-CNRS UMR 9221, France \& CEPII, France; email: \url{jerome.hericourt@univ-lille.fr}}\qquad Lise \textsc{Patureau}\thanks{Corresponding author.
Universit\'{e} Paris-Dauphine \& PSL University, LEDa, 75016 PARIS, FRANCE;  email: \url{lise.patureau@dauphine.fr} } }

\date{June 2018}
 \maketitle
\bigskip

\begin{abstract}
This paper investigates the pattern of international transport costs over time, using information contained in the US imports flows over 1974-2013. First, we document the importance of the per-unit (additive) component of transport costs. We thus find that additive costs are quantitatively sizable, representing between one third and one half of overall transport costs. Second, we identify the respective roles of the reduction in ``pure'' transport costs and trade composition effects in the downward trend of international transport costs, in the same spirit as \cite{hummels2007}. Unlike him, we find that trade composition effects do not matter much and, when they do, they tend to amplify (rather than reduce) the decrease in pure transport costs. Importantly, this difference of results can be attributed to the new way of modeling the per-unit component of transport costs we offer. In both aspects, our results point to the importance of the additive component in accounting for international transport costs.
\end{abstract}

@@

Trade costs are central to international economic analysis.
In particular, they are known to be a major obstacle to international economic integration and international trade flows.
Based on panel data, \citet{novy13} finds  that U.S.
trade costs with major trading partners declined on average by about 40 percent between 1970 and 2000.
Trade costs declined faster for more developed countries (\citealp{arvis2016trade}).
Still, several papers (mostly based on empirical estimates of the gravity equation) have shown that trade costs remain a major obstacle to trade (\citealp{Head_Mayer04} or \citealp{Disdier_Head08}, to name a few).
Using data over 1989-2000, \citet{anderson_wincoop_jel} thus estimate that average international trade costs represent a 74\% markup over production costs for industrialized countries.

Defined as the costs associated with the exchange of goods across national borders, trade costs are usually split between transaction costs (information costs, contract enforcement costs, costs associated with the use of different currencies...), policy costs (tariffs  and non-tariff costs), time costs (time to ship goods) and transport costs \emph{per se}.
In this vein, \citet{anderson_wincoop_jel} obtain that around 30\% of international trade costs are attributable to transport costs.
Equivalently, international transportation costs represent a 21\% markup over production costs.
Further, the sizable elasticity of trade with respect to freight costs obtained by \cite{Behar_Venables} (around -3) testifies to the impact of transport costs on trade flows.
If much of trade policy barriers have been removed over the second half of the twentieth century, this suggests that the transport costs component remains large and deserves attention.
International transport costs accordingly stand at the heart of the paper.\smallskip

This focus is shared with part of the international trade literature that studies the patterns of transport costs over time, such as \cite{hummels2007} and \cite{Behar_Venables}.
Many argue that transport costs have substantially decreased with technological advance in transportation (\citealp{levinson2016box}), infrastructure development and new communication technologies (\citealp{Lafourcade_Thisse}).
There is no consensus thought.
Distance as an incompressible cost is still a very important determinant of trade flows (\citealp{Disdier_Head08}); further, progress in transportation techniques (that reduces the constant-dollar cost of transport) should be compared with technical progress in production (that reduces the constant-dollar cost of production) to get a clear picture of the evolution of the transport costs burden.
In accordance with this view, \cite{hummels2007} points out that transport costs in proportion of export prices have declined more slowly than the common wisdom assumes.\smallskip

According to \cite{hummels2007}, this lower-than-expected decline of transport costs is partly explained by trade composition effects (i.e., changes over time in the baskets of imported goods and/or the origin countries).
Based on US imports over 1974-2004, he finds that total transport costs have decreased in both air and maritime shipment, but less so than ``pure'' transport costs (i.e., \textit{ceteris paribus} for given goods, mode, and trade routes) because trade composition effects have partly offset the pure transport costs decrease, thereby attenuating the overall transport costs downward trend.\footnote{Throughout the paper, we refer to ``pure'' or ``per se'' transport costs changes as the changes over time in the transport costs for a given product/country partner, ie excluding trade composition effects.} In this paper, we dig deeper into this issue.
As a major contribution, and in contrast to \cite{hummels2007}, we find that trade composition effects play a minor role in accounting for the trend patterns of transport costs.
Rather, the evolution of transport costs ``per se'' constitutes the main driving force behind the overall transport costs decrease, in particular for air transport.
In maritime shipment, trade composition effects matter more, but they amplify (rather than offset) the reduction in pure transport costs.
In this respect, we find that pure transport costs declined even less since the 1970s than argued by \cite{hummels2007}.\medskip

Proving this point drives us to provide a careful modelling of transport costs.
Following \citet{samuelson1954}, standard models of international trade have usually modeled trade costs as an \emph{ad-valorem} tax equivalent (ie, as a constant percentage of the producer price per unit traded, part of the ``iceberg cost'' hypothesis).\footnote{Strictly speaking, the ``iceberg cost'' includes both the ad-valorem dimension of trade cost and the fact that these costs are paid in terms of the good that is traded.
As this last element is irrelevant in our case, we use the terms ``iceberg'' and ``ad-valorem'' interchangeably, as it is commonly done in the literature.} However, a recent strand of the literature points out to the existence of an additive component, that is, a cost per unit traded (see, among others, \citealp{Irrazabal_2015} or \citealp{martin2012}).
Based on US sectoral data, our own estimates also bring out the importance of additive costs in international transport costs.
We obtain this by providing an empirical decomposition of the structure of transport costs over time, in which we explicitly distinguish between ad-valorem and additive costs.
To do so, we exploit information contained in the Census Bureau's database of US imports flows over 1974-2013.
This database contains a measure of international transport costs as the difference between the import and the export prices, on a year-product-partner country basis.
This database is the same as \cite{hummels2007}, extended over more recent years (2005-2013).
\smallskip

Our results may be summarized in two main new findings.
First, additive transport costs are quantitatively sizable.
On average over 1974-2013, the additive cost is estimated to be 1.8\% and 2.9\% of the export price in air and vessel transport respectively (at the 3 digit classification level). This is slightly lower than our estimates for the ad-valorem component (2.5\% and 3.2\% for air/vessel respectively on average over the period).
Put it differently, additive costs represent between one third and almost one half of the overall transport costs.
%\textbf{GD Pourquoi est-ce qu'en faisant le calcul je trouve 33\% et 47,5\% ?}
To the best of our knowledge, our paper is the first to provide such an extensive quantitative measure of both ad-valorem and additive costs in total transport costs.
This represents a valuable insight for calibrating related international trade or business cycle models.
We also provide an empirical assessment of what standard international trade models lose by skipping additive transport costs.
Quantitatively, the omission of the additive term leads to overestimate the ad-valorem component by roughly a factor~2. Further, the goodness of fit is systematically better when the additive component is included in the regression, even when taking into account the removed degree of freedom.\smallskip

Second, we find that the main source of the downward trend in overall transport costs -by roughly 50\% in air shipping and 60\% in vessel shipping over 1974-2013, comes mainly from the reduction of transport cost ``per se''.
Trade composition effects only have a limited influence, and when they matter (in maritime transport), composition effects amplify the reduction of pure transport costs.
Otherwise stated, pure transport costs have declined less than observed costs.
This stands in contrast to the conclusion obtained by \cite{hummels2007} on the same database (until 2004).
Our result thus deepens the paradox of the less-than-expected decline of transport costs.
Importantly, we show that the difference of results can be attributed to the new way of modeling of the additive component we offer.
Allowing the share of additive costs to vary across time, products and country partners (rather than being constant as in \citealp{hummels2007}), turns out to be key in the decomposition of underlying sources of the decrease of overall transport costs observed over the period.
\smallskip

In both aspects, our results point the importance of the additive component in accounting for international transport costs.
In this respect, it contributes to the literature that challenges the dominant role of iceberg costs in international trade (even in the case of ice transport, see \citealp{bosker2018ice}).
As shown by \cite{alchian}, the relative price of two varieties of some good will depend on the level of additive trade costs.
In this context, the relative demand for more expensive/higher quality product goods increases with trade costs (``shipping the good apples out'').
This is consistent with the findings by \citet{hummels_skiba}, who estimate the elasticity of freight rates with respect to price to be well below unity.
Also, their estimates imply that doubling freight costs increases average free alongside (fas) export prices by 80 to 141 percent, consistent with high quality goods being sold in markets with high freight costs.
Recent empirical studies based on micro-level data also provide a strong empirical support to the role of additive costs (i.e., cost per unit exported) in international trade costs.
Based on a firm-product-level database of French exporters, \citet{martin2012} finds that firms charge higher unit values for exports to more remote countries, supporting the importance of additive costs.
Using transaction-level trade data from Colombia, \cite{Lashkaripour-2017b} estimates that the transport cost elasticity varies greatly across industries, with a value of 0.8 for the average industry, in contrast with the iceberg specification.\footnote{Beyond the positive aspect, several recent papers also point out the normative implications of additive trade costs.
\citet{sorensen2014} extends \citet{melitz}'s seminal model of international trade by including additive trade costs, in addition to the ad-valorem component.
A key analytical result is that the welfare gain from a reduction in trade barriers is higher for a decrease in additive costs than a decrease in ad-valorem costs, due to the alteration of relative prices in a heterogeneous-firms trade framework.
This is confirmed by \cite{Irrazabal_2015}.
While these results suggest that important welfare gains can be achieved by reducing additive trade costs, not much progress has been done in quantifying such gains.
One potential reason is the lack of an empirical characterization of the additive component of trade costs (the one exception being \citealp{Irrazabal_2015}).
One contribution of the paper is to fill this gap.}

Closely related to our paper is the work by \citet{Irrazabal_2015}, which develops a structural framework for inferring relative additive trade costs from firm-level trade data.
Implementing their methodology on Norwegian firm-level export data for the year 2004, they find that, for the median shipment, additive costs (in Norwegian crowns) amount to 6\% of the export price multiplied by the ad-valorem cost (also expressed in Norwegian crowns).
Our results share in common with \citet{Irrazabal_2015} the important role of the additive component of international trade costs.
Yet, our paper complements their findings in two main respects.
First, while our study covers a narrower set of trade costs focused on transport costs, their data and their empirical approach only allows the identification of the ratio of the additive cost to the export price multiplied by the ad-valorem cost ; by contrast, our estimation strategy enables us to uncover separately both values of the ad-valorem and the additive costs.
From this, we can rebuild the ratio in similar terms to them, thereby gaining in generality in this respect.
While \citet{Irrazabal_2015} obtain that, for an export price multiplied by the ad-valorem cost of 100 USD, 6 USD are paid in additive trade costs, our own results point to a value paid in additive transport costs of 2.8 USD for maritime shipping and 1.5 USD for air transport for 2004.
This comparison of results suggests that between a third and a half of the additive trade costs are attributable to the transport sector (depending on the transport mode).
Second, we exploit exhaustive information about the imports flows of the US over a large time span from 1974 to 2013.
In this respect, our results deliver a broader view of the magnitude of additive costs in international trade over time.
In particular, we show that the modeling of additive costs is of critical importance in determining the underlying sources of the trends patterns of international transport costs observed since 1974.\footnote{On top of the previously cited papers about additive costs, our paper also relates to \citet{Kropf-Saure-JIE-2016}, which estimate the size and shape of per-shipment costs bases on Swiss data, and to \citet{Alessandria-et-al-AER-2010} or \citet{Hornok-et-al-JIE-2015, Hornok-et-al-RES-2015}, which point out the role of per-shipment costs (among which, administrative costs) in generating some ``lumpiness'' in international trade transactions.} \smallskip

@@

The paper is built as follows.
First, we estimate the values of international transport costs annually throughout the period 1974-2013 by transport mode, explicitly distinguishing between the additive and the ad-valorem components.
 Section \ref{sec:data_method} thus reports the mean values of the transport costs estimated over the period (by transport mode).
Relying on these estimation results, we then provide a decomposition of the transport costs trend patterns, between what comes from changes in the trade composition (by product and/or partner country), and what is attributable to changes in transport costs ``per se''.
Section \ref{sec:results_trends} is devoted to this analysis, while section \ref{sec:robustness} checks the robustness of our results to the separability assumption and the weighting scheme.
Section \ref{sec:conclu} concludes.

@@

Our analysis of transportation costs consists in exploiting the difference between commodity-level export and import prices, as in \cite{hummels2007}.
The database we use to construct our measure of transport costs comes from US annual "Imports of Merchandise" provided by the Census bureau, spanning from 1974 to 2013.
Details on the database are provided in Appendix \ref{app:data}.
We first use customs values, quantities and freight costs to recover free-alongside (fas) and cost-insurance-fret (cif) prices, by good, country of origin and transportation mode.\footnote{The related literature commonly refers to the fob price rather than the fas price.
The fas price (for ``Free Alongside Ship'') means that the seller must transport the goods all the way to the dock, close enough to be reached by the crane of the ship it will be transported in.
It is also the seller's responsibility to clear the goods for export.
The fob price (for ``Free on Board'') means that the seller is obligated to bring the goods all the way to the port, clear the goods for export, and see that they are loaded onto the ship nominated by the buyer.
Once the goods clear the railing of the vessel the buyer assumes the risk.
Note that this term is used exclusively for maritime and inland waterway transport.
While both terms are closely related, the US Foreign Trade Statistics reports fas prices.} More precisely, the (unit) fas price is computed as the total ``customs value'' in the US trade statistics divided by the shipping weight; in other words, it is the price for one kg of the good net of transportation costs.
The cif price is then computed as the sum of the customs value and freight charges, once again divided by the shipping weight.
Our dependant variable is finally computed as the ratio of the cif price divided by the fas price.
By construction higher than 1, the variable provides a measure of transport costs as a proportion of the good's price, an \emph{ad-valorem} equivalent.
This is a quite standard and widespread strategy, as emphasized by \citet{anderson_wincoop_jel}.

Obviously, this dataset has limitations.
First, it restricts our analysis to the study of international \emph{transport}  costs, as our measure of the cif-fas price gap only covers freight, insurance and handling costs.
It is thus silent about the others dimensions of international trade costs.
Second, in terms of transport costs \textit{per se}, our measure omits the ``indirect'' costs related to the time value of goods on their way to their export market (including holding cost for the goods in transit, inventory cost due to buffering the variability of delivery dates, preparation costs associated with shipment size...).
In this respect, this dataset embraces only a partial view of international transport barriers.
However, it is also clear that these direct transport costs do represent a sizable share of trade costs.
According to \citet{anderson_wincoop_jel}, the 21\% markup over production costs coming from transport costs includes both directly measured freight costs (11\%) and ``indirect'' costs that amount to a  9\% tax equivalent.
Further, the evidence summarized in \citet{anderson_wincoop_jel} points to a persisting importance of direct transport costs, especially compared to other trade barriers.
They remain more important than, e.g., policy barriers (8\% tax equivalent), language barrier (7\%) or information cost (6\%).\footnote{In his survey, \citet{Hummels_1999} mentions several papers which all point that transport costs pose a barrier similar in size, or larger than tariffs.
In the same vein, \citet{limao_venables} highlight the importance of infrastructures for trade costs in general, through their impact on transport costs.}

Exploiting this dataset also has three advantages.
First, it comes from a single, homogenous and trustworthy customs source.
This limits the measurement error bias.
Based on customs declarations, the US Imports Database inventories all imports (both values and quantities), by country of origin, to the United states at the HS 10-digit level, with a concordance code to the SITC 5-digit coding system.
%This will be crucial to compute transport costs, as we detail below.
In addition, the database reports information regarding freight, insurance and handling expenditures by transportation mode, land/ocean (or ``vessel'') and air.
We will make use of this to enlighten potential differences in the dynamics of transport costs across transportation mode.
Second, this dataset provides both the import price and the export price for the same good (for a given origin country).
This is highly valuable, as it gives us a direct measure of international transport costs (at the country/sector level), from which we can estimate separately the \textit{levels} of both the iceberg costs and of the additive costs, in contrast to e.g. \cite{Irrazabal_2015}.
Third, this dataset is available over a long time span, which we exploit for providing an analysis of the trend patterns of international transport costs over the period 1974-2013.\footnote{Note that, as mentioned by \cite{Lafourcade_Thisse}, our transport costs measures are based on actual trade flows, i.e.
ignoring those flows that did not happen because of presumably prohibitive transport costs.
In light of this, the estimated values that we obtain can be viewed as the lower bounds of the ``potential'' transport costs.} \smallskip

>>>>

\title{\textsc{International Transport costs:\\New Findings from modeling additive costs}\thanks{We are grateful to Holger Görg, S\'{e}bastien Jean and Vincent Vicard for their insightful comments. The paper has greatly benefited from remarks from the participants at ETSG 17\textsuperscript{th} (Helsinki), DEGIT XXII (Paris), seminars at the universities of California-Irvine, Paris-Dauphine, Paris-I and Nancy and the Kiel Institute for the World Economy.
Any remaining errors are ours. This paper was partly funded by a university grant from Lille-I University. It features an online appendix containing additional results, available on the authors' websites.}}

\author{Guillaume \textsc{Daudin}\thanks{%
University Paris-Dauphine \& PSL University, IRD, LEDa, UMR 225, DIAL, 75016 PARIS, FRANCE \newline
Sciences Po, Observatoire Fran\c{c}ais des Conjonctures \'{E}conomiques (OFCE), 75014 PARIS, FRANCE \newline
; email: \url{guillaume.daudin@dauphine.psl.eu}}  \qquad J\'{e}r\^{o}me \textsc{H\'{e}ricourt} \thanks{Universit\'{e} de Lille - LEM-CNRS UMR 9221, France \& CEPII, France; email: \url{jerome.hericourt@univ-lille.fr}}\qquad Lise \textsc{Patureau}\thanks{Corresponding author.
University Paris-Dauphine \& PSL University, LEDa, 75016 PARIS, FRANCE;  email: \url{lise.patureau@dauphine.psl.eu} } }

\date{December 2018}
 \maketitle
\bigskip

\begin{abstract}
This paper investigates the pattern of international transport costs over time, using information contained in the US imports flows from 1974 to 2013. First, we document the importance of the per-unit (additive) component of transport costs. We find that additive costs are quantitatively sizable, representing between one-third and one-half of overall transport costs. Second, we identify the respective roles of the reduction in ``pure'' transport costs and trade composition effects in the downward trend of international transport costs, in the same spirit as \cite{hummels2007}. Unlike him, we find that trade composition effects do not matter much and, when they do, they tend to amplify (rather than reduce) the decrease in pure transport costs. Importantly, this difference of results can be attributed to the new method of modeling the per-unit component of transport costs we offer. In both aspects, our results point to the importance of the additive component in accounting for international transport costs.
\end{abstract}

@@

Trade costs are central to international economic analysis.
In particular, they are known to be a major obstacle to international economic integration and international trade flows.
Based on panel data, \citet{novy13} finds  that US
trade costs with major trading partners declined on average by about 40\% between 1970 and 2000.
Trade costs declined faster for more developed countries (\citealp{arvis2016trade}).
Still, several papers (mostly based on empirical estimates of the gravity equation) have shown that trade costs remain a major obstacle to trade (\citealp{Head_Mayer04} or \citealp{Disdier_Head08}, to name a few).
Using data over 1989-2000, \citet{anderson_wincoop_jel} estimate that average international trade costs represent a 74\% markup over production costs for industrialized countries.

Defined as the costs associated with the exchange of goods across national borders, trade costs are usually split between transaction costs (information costs, contract enforcement costs, costs associated with the use of different currencies, etc.), policy costs (tariffs  and non-tariff costs), time costs (time to ship goods) and transport costs \emph{per se}.
In this vein, \citet{anderson_wincoop_jel} find that around 30\% of international trade costs are attributable to transport costs.
Equivalently, international transportation costs represent a 21\% markup over production costs.
Further, the sizable elasticity of trade with respect to freight costs obtained by \cite{Behar_Venables} (around -3) testifies to the impact of transport costs on trade flows. With many trade policy barriers removed over the second half of the 20th century, this suggests that the transport costs component remains large and deserves attention. International transport costs accordingly stand at the heart of this paper.\smallskip

This focus is shared with part of the international trade literature that studies the patterns of transport costs over time, such as \cite{hummels2007} and \cite{Behar_Venables}.
Many argue that transport costs have substantially decreased with technological advance in transportation (\citealp{levinson2016box}), infrastructure development and new communication technologies (\citealp{Lafourcade_Thisse}).
However, there is no consensus.
Distance as an incompressible cost is still a very important determinant of trade flows (\citealp{Disdier_Head08}); further, progress in transportation techniques (which reduces the constant-dollar cost of transport) should be compared with technical progress in production (which reduces the constant-dollar cost of production) to get a clear picture of the evolution of the transport costs burden.
In accordance with this view, \cite{hummels2007} points out that transport costs as a proportion of export prices have declined more slowly than is commonly assumed.\smallskip

According to \cite{hummels2007}, this lower-than-expected decline of transport costs is partly explained by trade composition effects (i.e. changes over time in the baskets of imported goods and/or the origin countries).
Based on US imports over 1974-2004, he finds that total transport costs have decreased in both air and maritime shipment, but less so than ``pure'' transport costs (i.e. \textit{ceteris paribus} for given goods, mode, and trade routes) because trade composition effects have partly offset the pure transport costs decrease, thereby attenuating the overall downward trend in transport costs.\footnote{Throughout the paper, we refer to ``pure'' or ``\textit{per se}'' transport costs changes as the changes over time in the transport costs for a given product/country partner, ie excluding trade composition effects.} In this paper, we dig deeper into this issue.
As a major contribution, and in contrast to \cite{hummels2007}, we find that trade composition effects play a minor role in accounting for the trend patterns of transport costs.
Rather, the evolution of transport costs \textit{per se} constitutes the main driving force behind the overall transport costs decrease, in particular for air transport.
In maritime shipment, trade composition effects matter more, but they amplify (rather than offset) the reduction in pure transport costs.
In this respect, we find that pure transport costs declined even less since the 1970s than argued by \cite{hummels2007}.\medskip

Proving this point drives us to provide a careful modeling of transport costs.
Following \citet{samuelson1954}, standard models of international trade have usually modeled trade costs as an ad-valorem tax equivalent (i.e. as a constant percentage of the producer price per unit traded, part of the ``iceberg cost'' hypothesis).\footnote{Strictly speaking, the ``iceberg cost'' includes both the ad-valorem dimension of trade cost and the fact that these costs are paid in terms of the good that is traded.
As this last element is irrelevant in our case, we use the terms ``iceberg'' and ``ad-valorem'' interchangeably, as is commonly done in the literature.} However, a recent strand of the literature points to the existence of an additive component, that is, a cost per unit traded (see, among others, \citealp{Irrazabal_2015}, or \citealp{martin2012}).
Based on US sectoral data, our own estimates also bring out the importance of additive costs in international transport costs.
We obtain this by providing an empirical decomposition of the structure of transport costs over time, in which we explicitly distinguish between ad-valorem and additive costs.
To do so, we exploit information contained in the Census Bureau's database of US imports flows over 1974-2013.
This database contains a measure of international transport costs as the difference between the import and the export prices, on a year-product-partner country basis.
This database is the same as that used by \cite{hummels2007}, extended over more recent years (2005-2013).
\smallskip

Our results may be summarized in two main new findings.
First, additive transport costs are quantitatively sizable.
On average over 1974-2013, the additive cost is estimated to be 1.8\% and 2.9\% of the export price in air and vessel transport respectively (at the 3 digit classification level). This is slightly lower than our estimates for the ad-valorem component (2.5\% and 3.2\% for air/vessel respectively on average over the period).
Put differently, additive costs represent between one-third and almost one-half of the overall transport costs.
%\textbf{GD Pourquoi est-ce qu'en faisant le calcul je trouve 33\% et 47,5\% ?}
To the best of our knowledge, our paper is the first to provide such an extensive quantitative measure of both ad-valorem and additive costs in total transport costs.
This represents a valuable insight for calibrating related international trade or business cycle models.
We also provide an empirical assessment of what standard international trade models lose by skipping additive transport costs.
Quantitatively, the omission of the additive term leads to overestimate the ad-valorem component by roughly a factor~2. Further, the goodness of fit is systematically better when the additive component is included in the regression, even when taking into account the removed degree of freedom.\smallskip

Second, we find that the main source of the downward trend in overall transport costs, by roughly 50\% in air shipping and 60\% in vessel shipping over 1974-2013, comes mainly from the reduction of transport cost \textit{per se}.
Trade composition effects have only limited influence, and when they matter (in maritime transport), composition effects amplify the reduction of pure transport costs.
Otherwise stated, pure transport costs have declined less than observed costs.
This stands in contrast to the conclusion obtained by \cite{hummels2007} on the same database (until 2004).
Our result thus deepens the paradox of the less-than-expected decline of transport costs.
Importantly, we show that the difference of results can be attributed to the new method of modeling the additive component that we offer.
Allowing the share of additive costs to vary across time, products and country partners (rather than being constant as in \citealp{hummels2007}), turn out to be key in the decomposition of underlying sources of the decrease of overall transport costs observed over the period.
\smallskip

In both aspects, our results point to the importance of the additive component in accounting for international transport costs.
In this respect, it contributes to the literature that challenges the dominant role of iceberg costs in international trade (even in the case of ice transport, see \citealp{bosker2018ice}).
As shown by \cite{alchian}, the relative price of two varieties of some goods will depend on the level of additive trade costs.
In this context, the relative demand for more expensive/higher-quality product goods increases with trade costs (``shipping the good apples out'').
This is consistent with the findings by \citet{hummels_skiba}, who estimate the elasticity of freight rates with respect to price to be well below unity.
Also, their estimates imply that doubling freight costs increases average free alongside ship (fas) export prices by 80\% to 141\%, consistent with high-quality goods being sold in markets with high freight costs.
Recent empirical studies based on micro-level data also provide strong empirical support for the role of additive costs (i.e. cost per unit exported) in international trade costs.
Based on a firm-product-level database of French exporters, \citet{martin2012} finds that firms charge higher unit values for exports to more remote countries, supporting the importance of additive costs.
Using transaction-level trade data from Colombia, \cite{Lashkaripour-2017b} estimates that the transport cost elasticity varies greatly across industries, with a value of 0.8 for the average industry, in contrast with the iceberg specification.\footnote{Beyond the positive aspect, several recent papers also point out the normative implications of additive trade costs. \citet{sorensen2014} extends \citet{melitz}'s seminal model of international trade by including additive trade costs, in addition to the ad-valorem component. A key analytical result is that the welfare gain from a reduction in trade barriers is higher for a decrease in additive costs than a decrease in ad-valorem costs, due to the alteration of relative prices in a heterogeneous-firms trade framework. This is confirmed by \cite{Irrazabal_2015}. While these results suggest that important welfare gains can be achieved by reducing additive trade costs, not much progress has been made in quantifying such gains. One potential reason is the lack of an empirical characterization of the additive component of trade costs (the one exception being \citealp{Irrazabal_2015}). One contribution of this paper is to fill that gap.}

Closely related to our paper is the work by \citet{Irrazabal_2015}, which develops a structural framework for inferring relative additive trade costs from firm-level trade data.
Implementing their methodology on Norwegian firm-level export data for the year 2004, they find that, for the median shipment, additive costs (in Norwegian crowns) amount to 6\% of the export price multiplied by the ad-valorem cost (also expressed in Norwegian crowns).
Our results share in common with \citet{Irrazabal_2015} the important role of the additive component of international trade costs.
Yet, our paper complements their findings in two main respects.
First, while our study covers a narrower set of trade costs focused on transport costs, their data and their empirical approach only allows the identification of the ratio of the additive cost to the export price multiplied by the ad-valorem cost; by contrast, our estimation strategy enables us to uncover separately both values of the ad-valorem and the additive costs.
From this, we can rebuild the ratio in similar terms to them, thereby gaining in generality in this respect.
While \citet{Irrazabal_2015} find that, for an export price multiplied by the ad-valorem cost of 100 USD, 6 USD are paid in additive trade costs, our results point to a value paid in additive transport costs of 2.8 USD for maritime shipping and 1.5 USD for air transport for 2004.
This comparison of results suggests that between a third and a half of the additive trade costs are attributable to the transport sector (depending on the transport mode).
Second, we exploit exhaustive information about the imports flows of the US over a wide time-span from 1974 to 2013.
In this respect, our results deliver a broader view of the magnitude of additive costs in international trade over time.
In particular, we show that the modeling of additive costs is of critical importance in determining the underlying sources of the trends patterns of international transport costs observed since 1974.\footnote{On top of the previously cited papers about additive costs, our paper also relates to \citet{Kropf-Saure-JIE-2016}, who estimate the size and shape of per-shipment costs based on Swiss data, and to \citet{Alessandria-et-al-AER-2010} or \citet{Hornok-et-al-JIE-2015, Hornok-et-al-RES-2015}, who point out the role of per-shipment costs (among which, administrative costs) in generating some ``lumpiness'' in international trade transactions.} \smallskip

@@

The paper is built as follows.
First, we estimate the values of international transport costs annually throughout the period 1974-2013 by transport mode, explicitly distinguishing between the additive and the ad-valorem components.
 Section \ref{sec:data_method} reports the mean values of the transport costs estimated over the period (by transport mode).
Relying on these estimation results, we then provide a decomposition of the transport costs trend patterns, between what arises from changes in the trade composition (by product and/or partner country), and what is attributable to changes in transport costs \textit{per se}.
Section \ref{sec:results_trends} is devoted to this analysis, while Section \ref{sec:robustness} checks the robustness of our results to the separability assumption and the weighting scheme.
Section \ref{sec:conclu} concludes.

@@

Our analysis of transportation costs consists of exploiting the difference between commodity-level export and import prices, as in \cite{hummels2007}. The database we use to construct our measure of transport costs comes from the US annual ``Imports of Merchandise'' provided by the US Census Bureau, spanning 1974 to 2013 (details on the database provided in Appendix \ref{app:data}). We first use customs values, quantities and freight costs to recover FAS and cost-insurance-freight (CIF) prices, by good, country of origin and transportation mode.\footnote{The related literature commonly refers to the FOB (free on board) price rather than the FAS price. The FAS price means that the seller must transport the goods all the way to the dock, close enough to be reached by the crane of the ship it will be transported in. It is also the seller's responsibility to clear the goods for export. The FOB price means that the seller is obligated to bring the goods all the way to the port, clear the goods for export, and see that they are loaded onto the ship nominated by the buyer. Once the goods clear the railing of the vessel, the buyer assumes the risk. Note that this term is used exclusively for maritime and inland waterway transport. While both terms are closely related, the US Foreign Trade Statistics reports FAS prices.} More precisely, the (unit) FAS price is computed as the total ``customs value'' in the US trade statistics divided by the shipping weight; in other words, it is the price for one kg of the goods net of transportation costs.
The CIF price is then computed as the sum of the customs value and freight charges, once again divided by the shipping weight.
Our dependant variable is finally computed as the ratio of the CIF price divided by the FAS price.
By construction higher than 1, the variable provides a measure of transport costs as a proportion of the goods' price, an ad-valorem equivalent.
This is quite a standard and widespread strategy, as emphasized by \citet{anderson_wincoop_jel}.

Obviously, this dataset has limitations.
First, it restricts our analysis to the study of international \emph{transport}  costs, as our measure of the CIF-FAS price gap only covers freight, insurance and handling costs.
It is thus silent about the other dimensions of international trade costs.
Second, in terms of transport costs \textit{per se}, our measure omits the ``indirect'' costs related to the time value of goods on their way to their export market (including holding cost for the goods in transit, inventory cost due to buffering the variability of delivery dates, preparation costs associated with shipment size, etc.).
In this respect, this dataset embraces only a partial view of international transport barriers.
However, it is also clear that these direct transport costs do represent a sizable share of trade costs.
According to \citet{anderson_wincoop_jel}, the 21\% markup over production costs coming from transport costs includes both directly measured freight costs (11\%) and ``indirect'' costs that amount to a  9\% tax equivalent.
Further, the evidence summarized in \citet{anderson_wincoop_jel} points to the persisting importance of direct transport costs, especially compared to other trade barriers.
They remain more important than, e.g., policy barriers (8\% tax equivalent), language barrier (7\%) or information cost (6\%).\footnote{In his survey, \citet{Hummels_1999} mentions several papers which all indicate that transport costs pose a barrier similar in size to, or larger than tariffs.
In the same vein, \citet{limao_venables} highlight the importance of infrastructures for trade costs in general, through their impact on transport costs.}

Exploiting this dataset also has three advantages.
First, it comes from a single, homogenous and trustworthy customs source.
This limits the measurement error bias.
Based on customs declarations, the US Imports Database inventories all imports (both values and quantities), by country of origin, to the United States at the HS 10-digit level, with a concordance code to the SITC 5-digit coding system.
%This will be crucial to compute transport costs, as we detail below.
In addition, the database reports information regarding freight, insurance and handling expenditures by transportation mode, land/ocean (or ``vessel'') and air.
We will make use of this to highlight potential differences in the dynamics of transport costs across transportation mode.
Second, this dataset provides both the import price and the export price for the same good (for a given origin country).
This is highly valuable, as it gives us a direct measure of international transport costs (at the country/sector level), from which we can estimate separately the \textit{levels} of both the iceberg costs and of the additive costs, in contrast to e.g. \cite{Irrazabal_2015}.
Third, this dataset is available over a long time span, which enables us to analyze the trend patterns of international transport costs over the period 1974-2013.\footnote{Note that, as mentioned by \cite{Lafourcade_Thisse}, our transport costs measures are based on actual trade flows, i.e.
ignoring those flows that did not happen because of presumably prohibitive transport costs.
In light of this, the estimated values that we obtain can be viewed as the lower bounds of the ``potential'' transport costs.} \smallskip

As detailed below, the use of a non-linear estimator triggers computational limitations that limit the level of possible detail, especially when covering a long period of time.
We estimate international transport costs at the 3-digit classification level, even if data series on the CIF and FAS prices are available at the 5-digit level.
However, we ensure the robustness of these results by conducting the estimations at the 4-digit level for some selected years.\footnote{The selected years for the 4-digit level estimations are: 1974, 1977, 1981, 1985, 1989, 1993, 1997, 2001, 2005, 2009, 2013.
Comparing different levels of aggregation is useful to check differences and the presence of biases precisely due to aggregation.
However, we obtain no substantial difference between the estimation results conducted at the 3 and 4-digit levels.
Estimation results at the 4-digit classification level are reported in Appendix \ref{app:4digit}.} Depending on the year, this leaves us with around 200 sectors at the 3-digit level, from around 200 countries of origin.

\subsection{Estimating transport costs: empirical specification}

\paragraph{The estimated equation} Our purpose is to provide time-varying estimates of the size of ad-valorem transport costs and additive transport costs.
To do so, we start from the equation that expresses the price $p$ (per kg) of goods paid by the importer (import, or CIF price) as a function of the producer price $\widetilde{p}$ (per kg) (export, or FAS price), given both additive (or per kg) transport costs ($t$) and ad-valorem ($\tau$) transport costs, according to:
